\section{Summation notation, product notation}

One should read 
\[\sum_{k = m}^na_m\] as the value of \verb|result| after executing this code:
\begin{verbatim}
    result = 0;
    for (k = m; k <= n; k = k + 1) {
        result = result + a_m.
    }
\end{verbatim}

Ideally, most of the time you should see the above sum as just a compactified way to write 
\[a_m + a_{m + 1} + \cdots + a_n\]

Notice that this code does not make sense if $m > n$, because the index $k$ is \textbf{always} incremented. In this case we always take the sum to be $0$ by default. So a sum like
\[\sum_{k = 1}^0k\] is equal to $0$, not $1$.

Similarly for products. one should read \[\prod_{k = m}^na_m\] as the value of \verb|result| after executing this code:

\begin{verbatim}
    result = 1;
    for (k = m; k <= n; k = k + 1) {
        result = result * a_m.
    }
\end{verbatim}

Again this code does not make sense if $m > n$. In this case we always take the product to be $1$ by default.

The following equations are some useful properties related to summations.
\[\sum_{i = m}^n(a_i + b_i) = \sum_{i = m}^na_i + \sum_{i = m}^nb_i\]
The above property is often colloquially referred to as ``splitting a sum''.
\[\sum_{i = m}^nca_i = c\sum_{i = 1}^ma_i\] if $c$ is a constant that does not depend on the index $i$.
\[\sum_{i = m}^na_i = a_m + \sum_{i = m + 1}^na_i\]
\[\sum_{i = m}^na_i = a_n + \sum_{i = n}^{n-1}a_i\]
Note that the above two equations make sense if $n \leq m$. (What happens when $n = m$?)

By yourself come up with some analogous formulas for products.

\section{The Principle of Mathematical Induction}

\begin{definition}
    A proposition $P(n)$ is a logical statement involving some value $n$.
\end{definition}

It is very helpful to consider an example or two. For example, define $P(n)$ to be the logical statement that $n = 1$. In other words, $P(n) \equiv (n = 1)$. Then $P(n) \equiv T$ if and only if $n = 1$, which is true but is silly to say. Another example: define $P(n) \equiv (2 \mid n)$. Then
\[P(n) \equiv \begin{cases}T, & 2 \mid n \\ F, & 2 \nmid n\end{cases}\]

But as of now we aren't really interested in examples like this. We'd like to know if there is a way to show that $P(n) \equiv T$ for $n \geq k$ for some integer $k$, for example.

For example, we'd like to figure out whether or not
\[\sum_{i = 1}^n(2i - 1) = n^2\] for all $n$. It seems true! But how might we prove it? This leads us to a technique called \textit{mathematical induction}, a powerful technique which lets us prove classes of statements indexed by the natural numbers. The technique of mathematical induction essentially relies on the following theorem. We will prove this theorem is true in a future section using a property of the natural numbers called the \textbf{Well-Ordering Principle}.

\begin{theorem}
Let $a$ be any natural number. And let $P(k)$ be a proposition about $k \in \mathbb{N}$. Assume the following claims are true:
\begin{itemize}
    \item $P(a) \equiv T$
    \item Assuming $P(k)$ is true for any natural number $k \geq a$, we can prove that $P(k + 1)$ is true.
\end{itemize}

Then $P(k)$ is true for all $k \geq a$.
\end{theorem}

Essentially, suppose you have an indexed family of propositions $P(n)$ which you want to prove true for $n \geq a$. A proof by mathematical induction is essentially the same as satisfying the hypotheses of the above theorem. Essentially, in a proof by induction you show the following:

\begin{enumerate}
    \item First we show that $P(a) \equiv T$.
    \item Next, assuming that $P(k)$ is true for $k \geq a$, prove that $P(k + 1)$ is true.
\end{enumerate}

Here is an example of this in action. We will prove the first equality stated at the beginning of the chapter.
\begin{proposition}
\[\sum_{i = 1}^n(2i - 1) = n^2\] for all $n$ for all natural numbers $n \geq 1$.
\end{proposition}
\begin{proof}
We will prove this equality using mathematical induction. First, we observe for $n = 1$ that
\[\sum_{i = 1}^1(2i - 1) = 1 = 1^2.\]
Now assume that 
\[\sum_{i = 1}^k(2i - 1) = k^2.\]
Then
\[\sum_{i = 1}^{k + 1}(2i - 1) = \sum_{i = 1}^k(2i - 1) + 2(k + 1) - 1 = k^2 + 2k + 1 = (k + 1)^2,\]
so we are done.
\end{proof}

Let's break down this proof a little bit further. Here the statement $P(n)$ is the statement that the equality 
\[\sum_{i = 1}^n(2i - 1) = n^2\] is true for that specific $n$. In the first part of the proof we show that $P(1)$ is true. In the second part of the proof we show that the truth of $P(k)$ implies the proof of $P(k + 1)$. To do this, we split the sum
\[\sum_{i = 1}^{k + 1}(2i - 1) = \sum_{i = 1}^k(2i - 1) + 2(k + 1) - 1\]
into two parts. By our hypothesis, the left part is exactly $k^2$. The rest of the prooof simplifies the sum to achieve the desired conclusion: that is to show 
\[\sum_{i = 1}^{k + 1}(2i - 1) = (k + 1)^2.\]

\section{Exercises}
\begin{enumerate}
    \item \input{Ch5/1_ex/problem_2}
    \item \input{Ch5/1_ex/problem_1}
    \item \input{Ch5/1_ex/problem_3}
    \item \input{Ch5/1_ex/problem_4}
    \item \input{Ch5/1_ex/problem_5}
    \item \input{Ch5/1_ex/problem_6}
\end{enumerate}

