\begin{itemize}
\item Show that the following sets are well-ordered under the ordering described. (It will be helpful to use the fact that $\mathbb{N}$ is well-ordered.
	\begin{itemize}
		\item $\mathbb{N} \cup \{\omega\}$, where $\omega > n$ for any $n \in \mathbb{N}$ and where $\mathbb{N}$ has the usual ordering.
		\item $\mathbb{N} \times \mathbb{N}$ under the dictionary order (which was defined in Example \ref{dict_order}). 
		\item $\mathbb{Z}$ under the following ordering $R$: $aRb$ if $|a| < |b|$, and if $|a| = |b|$ then $a < b$ if $a$ is negative and $b$ is positive.
	\end{itemize}
	
\item Show that the following sets are not well-ordered:
	\begin{itemize}
		\item The set of rationals $\mathbb{Q}$ under the usual ordering.
		\item The set $\{2^n \mid n \in \mathbb{Z}\}$ under the usual ordering on $\mathbb{Q}$.
		\item The cube $[0, 1] \times [0, 1]$ under the dictionary ordering.
	\end{itemize}
	
\item Show that for any well-ordered set $S$ with order relation $<$ and any element $r \in S$, either $r$ is a maximal element ($r \geq s$ for all $s \in S$) or $r$ has an immediate successor $r^{+}$, that is, a smallest element greater than $r$. Is every non-minimal element $r$ an immediate successor of some other element?

\item Suppose that $a$ is any integer and $b$ is any positive integer. State and prove when there exist integers $q$ and $r$ such that $r < |b/2|$ and we have 
\[a = qb + r.\]
When can we ensure that $q$ and $r$ are unique?

\item Prove that the Principle of Strong Induction implies the Principle of Mathematical Induction.

\item Prove the Principle of Strong Induction using the Well Ordering Principle.

\item Suppose that $P(n)$ is a proposition. In each part assume $P(n)$ satisfies the given list of properties and use the Well Ordering Principle to prove that $P(n)$ is true for all $n \geq 0$. After this, generalize.
	\begin{enumerate}
		\item 
		\begin{itemize}
			\item $P(0)$, $P(1)$, and $P(2)$ are true,
			\item $P(n) \equiv T \implies P(n + 3) \equiv T$.
		\end{itemize}
		\item
		\begin{itemize}
			\item $P(0)$ and $P(1)$ are true,
			\item If $P(n)$ is true, then $P(2n)$ is true,
			\item IF $P(n)$ is true, then $P(n - 1)$ is true.
		\end{itemize}
	\end{enumerate}
\end{itemize}